---
title: "Data Cleaning"
author: "Jacqueline Colao"
date: "12/9/2019"
output: html_document
---
---
title: "PS 239T Final Project - Data Cleaning"
author: "Jacqueline Colao"
date: "12/1/2019"
output: html_document
---

## 0) Set Up

### 0.1. Summary of File 

This file cleans the datasets gathered from the followthemoney API to ready them for data visualization. The main two things accomplished are creating a variable that is the percent of total political contributions given by an Interest Group in a year that are given to party committees and creating a dataset that can be used to create a stacked bar graph showing the total political contributions made by an interest group in a year broken down by the amount given to candidates and the amount given to party committees. 

### 0.2. Setup Environment

```{r}
# remove all objects
rm(list=ls())

# check working directory
getwd()

# Unload all packages 
library(pacman)
pacman::p_unload(all)

# Add packages 
pacman::p_load(
  tidyverse, #dplyr, readr, etc.
  foreign, #load data types including stata .dta files 
  magrittr, #%<>% operator
  skimr, #for summerising
  readr, #to read file from git repository
  httr
)

```

### 0.3. Load dataset

```{r}
candidate_data <-read.csv("https://raw.github.com/jacquelinecolao/ps239t-final-project/master/Data/Raw%20Data/Interest%20Group%20Candidate%20Donations.csv")

partyCommittee_data<-read.csv("https://raw.github.com/jacquelinecolao/ps239t-final-project/master/Data/Raw%20Data/Interest%20Group%20Party%20Committee%20Donations.csv")

totalContributions_byYear_data<-read.csv("https://raw.github.com/jacquelinecolao/ps239t-final-project/master/Data/Raw%20Data/Interest%20Group%20Total%20Contributions.csv")
```

## 1) Summarise Data 

### 1.1. Totals for Party Committee Data 

```{r}
#get total contributions to party committees by each Interest Group for each year
partyCommittee_data %<>%
  #group the data by Interest Group and Year
  group_by(Interest.Group,Year) %>%
  #sum the contributions for the Interest Group in the Year
  mutate(total_contributions_IC_Year=sum(Amount.Contributed)) %>%
  #ungroup so that each party committee donation also has a variable with the total party committee donations
  ungroup()
```

### 1.2. Totals for Candidate Data 

```{r}
#get total contributions to candidate by an Interest Group for each year
#same process as above 
candidate_data %<>%
  group_by(Interest.Group,Year) %>%
  mutate(total_contributions_IC_Year=sum(Contribution)) %>%
  ungroup()
```

## 2) Arrange Data

### 2.1. Arrange Party Committee Data 

```{r}
#arranges the data alphabetically by Interest Group - showing the most recent year first for each Interest Group
partyCommittee_data %<>%
  arrange(Interest.Group,desc(Year)) 
```

### 2.2. Arrange Candidate Contribution Data 

```{r}
#arranges the candidate data in the same way as the party committee data
candidate_data %<>%
  arrange(Interest.Group,desc(Year))
```


### 2.3. Arrange Total Contributions Data 

```{r}
#same for total contribution data
totalContributions_byYear_data %<>% 
  arrange(Interest.Group,desc(Year))
```

## 3) Getting Data Ready to Graph - Creating New Variables 

### 3.1. Preparation for Graphing the Percent of Total Contributions Given to Party Committees 

```{r}
#merge the total contributions data with the party committee data 
total_contributions_plus_partyCommittee=left_join(totalContributions_byYear_data,partyCommittee_data,by=c("Interest.Group","Year"))

#replace NAs with 0s
total_contributions_plus_partyCommittee %<>%
  mutate_all(funs(replace(., is.na(.), 0))) %>%
  #rename variables 
  rename("Total_Contributions" = Amount.Contributed.x) %>%
  rename("Party_Committee_Contribution" = Amount.Contributed.y) %>%
  rename("Total_Party_Committee_Contribution_by_Year" = total_contributions_IC_Year) %>%
  #create a new variable that is the percent of total contributions given to party committees
  mutate("Percent_Donated_to_Party_Committees" = Total_Party_Committee_Contribution_by_Year/Total_Contributions)

write.csv(total_contributions_plus_partyCommittee,"/Users/jacquelinecolao/Desktop/ps239t-final-project2/Data/Cleaned Data for Graphs/Party Committee Data with Total Contribution Variable.csv")
```

### 3.2. Preparation for Graphing a Stacked Bar Graph of party committee donations and candidate donations 

```{r}
#need to get the data in a form to be able to make a stacked bar graph
total_contributions_plus_partyCommittee %<>%
  #create a variable that is the amount the Interest Group contributed to candidates
  mutate(amount_to_candidate = Total_Contributions-Total_Party_Committee_Contribution_by_Year) %>%
  #create a categorical variable of whether the amount contributed is to a candidate or party committee - here it is to party committees 
  mutate(recipient = "Party Committee")
  

#copy the above data frame 
data_2<-total_contributions_plus_partyCommittee
#create a categorical variable that indicates that the amount contributed is to a candidate 
data_2 %<>%
  mutate(recipient = "Candidate") %>%
  #delete the party committee contribution variable 
  select(-c(Total_Party_Committee_Contribution_by_Year)) %>%
  #rename candidate contribution variable to amount_contributed 
  rename(amount_contributed = amount_to_candidate) %>%
  #remane total contribution variable 
  rename(total_contributions = Total_Contributions)

total_contributions_plus_partyCommittee %<>%
  #delete to candidate contribution variable 
  select(-c(amount_to_candidate)) %>%
  #rename party contribution variable to amount_contributed 
  rename(amount_contributed = Total_Party_Committee_Contribution_by_Year) %>%
  #remane total contribution variable 
  rename(total_contributions = Total_Contributions)

#combine the two datasets 
data <- bind_rows(total_contributions_plus_partyCommittee,data_2)

#arrange
data %<>% arrange(Interest.Group,desc(Year))

#get rid of the party committees and donations to specific committees
data %<>%
  select(-c(Party.Committee,Party_Committee_Contribution,Percent_Donated_to_Party_Committees)) %>%
  #delete duplicate rows
  distinct(Interest.Group, Year,recipient, .keep_all = TRUE)


write.csv(data,"/Users/jacquelinecolao/Desktop/ps239t-final-project2/Data/Cleaned Data for Graphs/Candidate and Party Contribution Data - Stacked Graph.csv" )

```